---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

For this task, we'll be using a unz connection. So whoever wants to reproduce this analysis doesn't have to carry about uncompressing the file.

```{r settingup, results=FALSE, message=FALSE}

setwd("/Users/fabianoal/Documents/Coursera/Repos/RepData_PeerAssessment1")
packages <- c("data.table", "dplyr", "lattice", "lubridate", "RColorBrewer")
sapply(packages, require, character.only=TRUE, quietly=TRUE)
fileName <- "./activity.zip"
f <- read.csv(unz(fileName, "activity.csv", encoding = getOption("encoding")), stringsAsFactors=FALSE)
f$date <- as.Date(f$date)
```

## What is mean total number of steps taken per day?

We'll be using the dplyr to calculate the amount of steps per day, and them, plot a histogram.

```{r firstanalysis}
f.totalsteps <- f %>%
  filter(complete.cases(.)) %>%
  group_by(date) %>%
  summarise(amountSteps = sum(steps))

barplot(f.totalsteps$amountSteps, names=f.totalsteps$date, xlab="Amount of Steps", col="#1d91c0", main="Amount of Steps per Date")
```

The mean of steps per day is `r format(mean(f.totalsteps$amountSteps), digits=0, scientific=FALSE)` and the median is `r median(f.totalsteps$amountSteps)`.

## What is the average daily activity pattern?

Now, we summarise the data not by date, but by interval so we can see the activity daily pattern.

```{r secondanalysis}
f.actvPattern <- f %>%
  filter(complete.cases(.)) %>%
  group_by(interval) %>%
  summarise(meanSteps = mean(steps))


xyplot(f.actvPattern$meanSteps~f.actvPattern$interval, ylab="Average Amount of Steps", xlab="Interval", type="b", col="#1d91c0", main="Daily Activity Pattern")

```

The 5-minute interval that, in average, has the highest number of steps is the `r arrange(f.actvPattern, desc(meanSteps))[1, "interval"]`th.

## Imputing missing values

The original dataset has **`r sum(is.na(f$steps))`** NAs values. Thus, for filling up the missing values, we will use the average number of steps for that specific interval.

To do this, we can use the *f.actvPattern* data frame created on the *secondanalysis* chunk.

```{r thirdanalysis, results="hide"}

ff <- merge(f,f.actvPattern, all.x = TRUE) %>%
  mutate(filledSteps = ifelse (is.na(steps), meanSteps, steps))

```

Ok. Now, let's do another histogram to see if the introduction of values for the NA observations has caused some changes...


```{r fourthanalysis_part_two}
ff.totalsteps <- ff %>%
  group_by(date) %>%
  summarise(filledAmountSteps = trunc(sum(filledSteps)), amountsteps=sum(steps))

barplot(ff.totalsteps$filledAmountSteps, names=ff.totalsteps$date, xlab="Amount of Steps", col="#1d91c0", main="Amount of Steps per Day - Filled NAs")
```

Now, let's see how the values has changed with the filling of NAs values.

* Mean of steps from filled dataset: `r format(mean(ff.totalsteps$filledAmountSteps), digits=0, scientific = FALSE)`
* Mean of steps from original dataset: `r format(mean(f.totalsteps$amountSteps), digits=0, scientific = FALSE)`
* Median of steps from filled dataset: `r format(median(ff.totalsteps$filledAmountSteps),digits=0, scientific = FALSE)`
* Median of steps from original dataset: `r format(median(f.totalsteps$amountSteps),digits=0, scientific = FALSE)`

## Are there differences in activity patterns between weekdays and weekends?

Now, for showing the differences between weekdays and weekends, we create a new variable *tweekday* on the filled data set which consists of a factor with two levels, then we take the average steps per *interval* and *tweekday* so we can plot the timeseries showing de different behaviors.

```{r fifithanalysis}
ff$tweekday <- factor(ifelse(wday(ff$date) %in% 1:5, 'weekday','weekend'))

ff.actvPattern <- ff %>%
    group_by(tweekday, interval) %>%
    summarise(meanSteps = mean(filledSteps))
trellis.par.set(strip.background=list(col="#7fcdbb"))
xyplot(meanSteps ~ interval | tweekday, data=ff.actvPattern, ylab="Average Number of Steps", xlab="Interval", type="b", col="#1d91c0", main="Daily Activity Pattern", layout = c(1,2))
```
